% --- ACRONYMS ---

\newacronym[
  description={Atmospheric Boundary Layer. The lowest part of Earth’s atmosphere, directly influenced by the Earth’s surface}
]{abl}{ABL}{Atmospheric Boundary Layer}

\newacronym[
  description={Boundary Representation. File format to store 3D models that consists of vertices, edges, wires, faces, shells, solids, etc.}
]{brep}{BRep}{Boundary Representation}

\newacronym[
  description={Utility to calculate the Leaf Area Index, used during the urbanMicroclimateFoam solver execution}
]{calclai}{calcLAI}{Leaf Area Index Calculator}

\newacronym[
  description={Computational Fluid Dynamics}
]{cfd}{CFD}{Computational Fluid Dynamics}

\newacronym[
  description={Digital Elevation Model. Digital representation of terrain elevation that consists of a regular grid of elevation z-values}
]{dem}{DEM}{Digital Elevation Model}

\newacronym[
  description={Standard weather data file used by EnergyPlus simulations, containing hourly or sub-hourly weather data for a specific location}
]{epw}{EPW}{EnergyPlus Weather}

\newacronym[
  description={OpenFOAM function that groups fine mesh faces into coarse mesh faces to reduce computational time}
]{faceagglomerate}{faceAgglomerate}{Face Agglomeration}

\newacronym[
  description={Leaf Area Density. Total one-sided leaf area per unit volume within a plant canopy}
]{lad}{LAD}{Leaf Area Density}

\newacronym[
  description={Leaf Area Index. Total one-sided leaf area per unit of ground area within a plant canopy}
]{lai}{LAI}{Leaf Area Index}

\newacronym[
  description={Level of detail used to represent 3D models with varying geometric complexity}
]{lod}{LOD}{Level of Detail}

\newacronym[
  description={Mean Bias Error. Statistical metric that quantifies the average deviation between predicted and observed values}
]{mbe}{MBE}{Mean Bias Error}

\newacronym[
  description={Mean Squared Error. Statistical metric that measures the average squared difference between predicted and observed values}
]{mse}{MSE}{Mean Squared Error}

\newacronym[
  description={High-performance multi-physics finite-element software}
]{netgen}{NETGEN}{Netgen}

\newacronym[
  description={Open-source field operation and manipulation software for CFD simulations}
]{openfoam}{OpenFOAM}{Open Field Operation And Manipulation}

\newacronym[
  description={ReynoldsAveraged Navier-Stokes. Time-averaged equations of motion for fluid flow}
]{rans}{RANS}{Reynolds-Averaged Navier-Stokes}

\newacronym[
  description={Root Mean Squared Error. Statistical metric that measures the average difference between predicted and observed values}
]{rmse}{RMSE}{Root Mean Squared Error}

\newacronym[
  description={Mean Absolute Error. Statistical metric that measures the average measures the average magnitude of error}
]{mae}{MAE}{Mean Absolute Error}

\newacronym[
  description={OpenFOAM utility to generate a 3D mesh containing hexahedra automatically from triangulated surface geometries}
]{snappyhexmesh}{snappyHexMesh}{snappyHexMesh}

\newacronym[
  description={Utility to identify which boundary faces are directly illuminated by the sun for each sun position at time steps}
]{solarraytracinggen}{solarRayTracingGen}{Solar Ray Tracing Generator}

\newacronym[
  description={Standardized 3D model file format used for exchanging product data between different CAD software}
]{step}{STEP}{Standard for the Exchange of Product Data}

\newacronym[
  description={Standardized 3D model file format used to represent geometry as a collection of triangular facets}
]{stl}{STL}{Stereolithography}

\newacronym[
  description={Mean Radiant Temperature. Average temperature of all surfaces surrounding a person}
]{mrt}{MRT}{Mean Radiant Temperature}

\newacronym[
  description={Typical Meteorological Year — representative set of hourly meteorological data for a location}
]{tmy}{TMY}{Typical Meteorological Year}

\newacronym[
  description={Urban Heat Island. Phenomenon where urban areas are hotter than rural areas due to materials and human activities}
]{uhi}{UHI}{Urban Heat Island}

\newacronym[
  description={OpenFOAM utility to calculate view factors from face-agglomeration arrays}
]{viewfactors}{viewFactorsGen}{viewFactors Generator}

\newacronym[
  description={Universal Thermal Climate Index. Physiological comfort of human body index}
]{utci}{UTCI}{Universal Thermal Climate Index}

\glsaddall
\makeListOfAcronyms
