\begin{acknowledgments}

I would like to express my deepest appreciation to my five committee members: Dr. Patrick Kastner, Dr. Athanassios Economou, Tyrone Marshall, Dr. Paula Gomez, and Dr. Marcelo Bernal for their support, consideration, insightful comments and observations. I'm extremely grateful to Professor Patrick Kastner, my thesis advisor, for his unparalleled knowledge and expertise, enthusiasm, motivation, and all the incredible support provided through my graduate studies. I started with him and continued with him throughout my graduate studies. His VIP Surrogate Models for Urban Regeneration was definitely what introduced me to the field of Urban Microclimate Modeling, which is the field that I want to become an expert in. The way he works is something I want to replicate. He is extremely rigorous and professional. He manages a lot of online repositories and he keeps track of everything simultaneously. I truly value all the instances that he generated to help us with collaborate with other disciplines and institutions.

As another member of the committee, I would also like to extend my deepest gratitude to Dr. Paula Gomez, who has trusted in my work, granting me opportunities I could have never imagined. She brought me here and provided me with the opportunity to study and pursue my degree. She singledhandly changed my pessimistic worldview into one of possibilities and that I can accomplish and dream for more than I thought. Dr. Gomez allowed me to work as a Graduate Research Assistance at the Georgia Tech Research Institute under her guidance, granting me exceptional research, technical, and soft skills that I highly value. Additionally, I'm deeply indebted to Dr. Marcelo Bernal, who always supported my career choices and constantly motivated me to continuously grow. Dr. Bernal has had a long-standing influence on my life from the time he was my instructor back in Chile and then became here to pursue his doctorate. He was the first academic advisor that I had. I am an architect by training and a computational designer by choice thanks to him. He provided me with a summer internship with at Perkins\&Will and was one of the best opportunities I could ask for. Along with Dr. Gomez, Dr. Bernal trusted me and opened many doors without any doubt, allowing me to share what I'm capable of. 

During my time at Georgia Tech and Perkins\&Will I had the opportunity of meeting Tyrone Marshall. I would like to extend my sincere thanks him, who was instrumental in defining my specific field of interest in computational design. His Building Simulation in Design Practice class was an amazing platform to delve deeper in climate, serving as one of the main drivers of this work. He also supported me during my internship period at Perkins\&Will, allowing me to showcase my work to other designer colleagues which I truly appreciate. I am also grateful to Dr. Athanassios (Thanos) Economou who undoubtedly supported me, again, trusting and valuing the quality of my work. I am glad that I joined the design computation program under his guidance. I will never forget the incredible enthusiasm he showed the first time I met him when I joined the Design Computation program. I had the opportunity to attend his Shape Grammars class, which resulted in one of the best academic experiences I have ever had. Dr. Economou invited me to review his architecture studio class, further enriching my academic experience and allowing me to be part of other students' academic formation.

The completion of this work would not have been possible without the incredible assistance of Robin Tucker. Her role as the Academic Advisor in the School of Architecture is what makes all programs work seamlessly. I highly appreciate her promptness, dedication, and consideration from the time I was accepted at Georgia Tech as a graduate student until the moment I finished my studies. Special thanks to the Georgia Tech Research Institute for allowing me to work as a Graduate Research Assistant during my studies. The support and flexibility they provided was crucial in my career development and successfully completing my graduate studies.

\clearpage
None of this work could have been properly done without the unconditional support of my dearest friend Suilim. Her companionship has been undoubtedly essential during the last years of my life. I do not have the words to express my gratitude to my friend Gonzalo for his unwavering support, advice, teaching, and enormously facilitating my life here at Georgia Tech. His company is what made my adventure at Georgia Tech extremely valuable and fun.

Finally, I'm extremely grateful to my mother Rosa, my brother Nicolas, my stepmother Marcela, my uncle Jose, my cousins Jimena and Jose, and my stepfather Jose Luis. My relatively small family is everything that I need. Their constant and unconditional support is what allowed me to leave my home country and study abroad. I truly appreciate them for being so understanding for my absence, limited time, and being so proud and happy about every small success that brought me here.

\end{acknowledgments}
